---
title: "Test_phylip_trees"
author: "Matthew Bootsma"
date: "October 29, 2018"
output: html_document
---
This script builds a tree in the package poppr
it then plots the tree in ape
still need to figure out how to export the tree object
```{r setup, include=FALSE}
library("adegenet")
library("poppr")
library("ape") # To visualize the tree using the "nj" function
library("geiger")
library("magrittr")
library("tictoc")
library("pegas")
library("devtools")
library("utils")
library(phylotools)
library("dendextend", lib.loc="~/R/win-library/3.5")
```


#Load Data
```{r}
WAE_data = read.genepop("../SNPS/v6_keep_max_maf_byLocus.recode.gen", ncode = 3)
```

#Trees
  NOTE: use the "cutoff = " argument in "aboot()" function to set the minimum bootstrap value to report.
```{r ,build tree (10 individuals)}
WAE_genpop_nei = genind2genpop(WAE_data, pop = WAE_data@pop) %>% 
  aboot(cutoff = 0, quiet = TRUE, sample = 1000, distance = nei.dist)

plot(WAE_genpop_nei)
#format names to remove sample ID
WAE_genpop_nei$tip.label = gsub("\\d+","",WAE_genpop_nei$tip.label)
WAE_genpop_nei$tip.label = gsub("-","",WAE_genpop_nei$tip.label)
#remove decimal precision
WAE_genpop_nei$node.label = round(WAE_genpop_nei$node.label,0)
#set colors based on repU
WAE_genpop_nei$tip.label

repu_mono_MN_shades = c("#FF9933","#0000FF","#0000FF",
         "#9933FF","#00cc99","#9933FF",
         "#000000","#0000FF","#00cc99",
         "#000000","#9933FF","#0000FF",
         "#00cc99","#00cc99","#666666",
         "#00cc99","#00cc99","#000000",
         "#00cc99","#99cc00","#9933FF",
         "#0000FF","#FF6633")


repu_unique_MN_shades = c("#FF9933","#0000FF","#0000FF",
         "#9933FF","#006666","#9933FF",
         "#000000","#0000FF","#003333",
         "#000000","#9933FF","#0000FF",
         "#00cc99","#009966","#666666",
         "#006633","#339933","#000000",
         "#33ff00","#99cc00","#9933FF",
         "#0000FF","#FF6633")
```

#read tree 
If you've already made one and just want to work with the graphical representation, start here.

NOTE: This will CHANGE THE ORDER of your tip.label, so consider this when you are trying to fit colors to tips
  ALSO, consider that other things could be reorganized as well. I have yet to fully investigate this.
```{r}
imported_tree = read.tree("./tree_from_v6_pop_gen_snps.tree")
imported_tree$tip.label
```
#plot tree
Many options available, see ape documentation
```{r}
pdf("./tree.pdf", height = 24, width = 18.54545)
plot(imported_tree, type = "phylogram",
     use.edge.length = TRUE,
     node.pos = NULL, 
     show.tip.label = TRUE, 
     show.node.label = F,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0.0005, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     tip.color = repu_mono_MN_shades, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
nodelabels(imported_tree$node.label, adj = c(1.3, -0.5), frame = "n", cex = 1.5,
           font = 3, xpd = TRUE)
axisPhylo(3)
plot(imported_tree, type = "phylogram",
     use.edge.length = TRUE,
     node.pos = 2, 
     show.tip.label = TRUE, 
     show.node.label = T,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     tip.color = repu_mono_MN_shades, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
dev.off()
```
#Save tree as object
```{r}
write.tree(WAE_genpop_nei, "./tree_from_v6_pop_gen_snps.tree")
```


```{r}

WAE_tree_reorder = reorder.phylo(WAE_genpop_nei, order = "cladewise", index.only = F)

WAE_genpop_nei$edge
WAE_tree_reorder$edge

plot(WAE_genpop_nei)
plot(WAE_tree_reorder)
```

#test rotate
```{r, read tree}
WAE_genpop_nei = read.tree("./tree_from_v6_pop_gen_snps.tree")
```

```{r, plot tree using reorder}
repu_mono_MN_shades2 = c("#00cc99","#996600","#0000FF",
         "#FF9900","#0000FF","#ffcc00",
         "#000000","#0000FF","#0000FF",
         "#0000FF","#000000","#000000",
         "#9933FF","#9933FF","#666666",
         "#9933FF","#9933FF","#00cc99",
         "#00cc99","#00cc99","#00cc99",
         "#00cc99","#00cc99")


pdf("./v6_natural_pops_rotate_tree.pdf", height = 24, width = 18.54545)
WAE_genpop_nei %>% 
  rotateConstr(constraint = c(
    #other
    "St_Louis_River","WolfR_",
    "Pike_River",
    "Manitowish_Lake","Eau_Claire_River",
    "Lake_Millicent","Chippewa_Flowage","Turtle_Flambeau_Flowage",
    "Sanford_Lake",
    "Escanaba","Big_Arbor_Vitae",
    "Willow_Flowage","Kawaguesaga","Delavan_",
    "Medicine_Lake",
    "Lake_Wisconsin_",
    "Mille_Lacs","Cutfoot_Sioux","Sarah_Lake","Lake_Koronis","Ottertail_Lake","Pine_River","Red_Lake")
    ) %>% 
  plot(type = "phylogram",
     tip.color = repu_mono_MN_shades2, 
     use.edge.length = TRUE,
     node.pos = 1, 
     show.tip.label = TRUE, 
     show.node.label = T,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
dev.off()
```


