---
title: "Test_phylip_trees"
author: "Matthew Bootsma"
date: "October 29, 2018"
output: html_document
---
NOTE: exported tree can allegedly be formatted in "mega" which is a good GUI for tree making, have yet to investigate but for future work I'll be investigating this.


This script builds a tree in the package poppr
it then plots the tree in ape
still need to figure out how to export the tree object

NOTE: The package geiger has some weird interaction issues with other packages. If it doesn't want to load just restart R and then go through this analysis before running other scripts.
```{r setup, include=FALSE}
library("adegenet")
library("poppr")
library("ape") # To visualize the tree using the "nj" function
library("geiger")
library("magrittr")
library("tictoc")
library("pegas")
library("devtools")
library("utils")
library(phylotools)
library("dendextend", lib.loc="~/R/win-library/3.5")
```


#Load Data
```{r}
WAE_data = read.genepop("../SNPS/v7_IAGLR_snps.recode.gen", ncode = 3)
```

#Trees
  NOTE: use the "cutoff = " argument in "aboot()" function to set the minimum bootstrap value to report.
```{r ,build tree (10 individuals)}
WAE_genpop_nei = genind2genpop(WAE_data, pop = WAE_data@pop) %>% 
  aboot(cutoff = 75, quiet = TRUE, sample = 1000, distance = nei.dist)

plot(WAE_genpop_nei)
#format names to remove sample ID
WAE_genpop_nei$tip.label = gsub("\\d+","",WAE_genpop_nei$tip.label)
WAE_genpop_nei$tip.label = gsub("-","",WAE_genpop_nei$tip.label)
#remove decimal precision
WAE_genpop_nei$node.label = round(WAE_genpop_nei$node.label,0)
#set colors based on repU
WAE_genpop_nei$tip.label


repu_unique_MN_shades = c("#FF9933","#0000FF","#0000FF",
         "#9933FF","#006666","#9933FF",
         "#000000","#0000FF","#003333",
         "#000000","#9933FF","#0000FF",
         "#00cc99","#009966","#666666",
         "#006633","#339933","#000000",
         "#33ff00","#99cc00","#9933FF",
         "#0000FF","#FF6633")
```


#Save tree as object
```{r}
write.tree(WAE_genpop_nei, "./tree_from_v6_pop_gen_snps.tree")
```

#read tree 
If you've already made one and just want to work with the graphical representation, start here.

NOTE: This will CHANGE THE ORDER of your tip.label, so consider this when you are trying to fit colors to tips
  ALSO, consider that other things could be reorganized as well. I have yet to fully investigate this.
  
```{r}
WAE_genpop_nei = read.tree("./tree_from_v6_pop_gen_snps.tree")
WAE_genpop_nei$tip.label
```

#plot tree (Base R)
Many options available, see ape documentation
OR
see ggplot section below (rotate nodes and tips developed here)
```{r}
pdf("./tree.pdf", height = 24, width = 18.54545)
plot(WAE_genpop_nei, type = "phylogram",
     use.edge.length = TRUE,
     node.pos = NULL, 
     show.tip.label = TRUE, 
     show.node.label = F,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0.0005, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     tip.color = repu_mono_MN_shades, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
nodelabels(WAE_genpop_nei$node.label, adj = c(1.3, -0.5), frame = "n", cex = 1.5,
           font = 3, xpd = TRUE)
axisPhylo(3)
plot(WAE_genpop_nei, type = "phylogram",
     use.edge.length = TRUE,
     node.pos = 2, 
     show.tip.label = TRUE, 
     show.node.label = T,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     tip.color = repu_mono_MN_shades, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
dev.off()
```



# Develop rotate method

This will require some back and forth of 1) identify nodes to flip, 2) flip nodes 3) constrain tip order
The first plot code will give you ID's of tips and nodes
The second plot will show you how changes you apply are formatted

IF YOU FIND A BETTER WAY, please contact me at boots17917@gmail to let me know and I will update this repository
thanks!
```{r, plot tree using reorder}
#0000FF --- WISC
#9933FF --- CHIP
#
#initialize label colors based on tip.names order
repu_mono_MN_shades2 = c("#cc8b00","#0000FF",
                         "#0000FF","#9933FF",
                         "#000000","#9933FF",
                         "#0000FF","#000000",
                         "#9933FF","#9933FF",
                         "#0000FF","#000000",
                         "#000000","#000000",
                         "#000000","#000000",
                         "#0000FF","#000000",
                         "#000000","#9933FF",
                         "#0000FF","#cc5f00")
#Rotate one node at a time, start at tips and work back up the tree
#I tried a vector of the nodes but it didn't give me the results I expected
#This does, though I did have to combine this approach with the rotateConstr function in my plot
#If I didn't do that, then there were lines intersecting.




# WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei_rotate, node = 42)
# WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei_rotate, node = 41)

#This is a base plot that will give you the ID of node and label numbers to rotate
pdf("./IAGLR_snps_rotate_tree.pdf", height = 24, width = 18.54545)
WAE_genpop_nei %>% 
  # rotateConstr(constraint = c(
  #   "Medicine_Lake",
  #   "Lake_Wisconsin",
  #   "WolfR_"
  #   
  #   
  #   # "Escanaba",
  #   # "Willow_Flowage",
  #   # "Big_Arbor_Vitae",
  #   # "Manitowish_Lake",
  #   # "Lake_Millicent",
  #   # "Chippewa_Flowage",
  #   # "Turtle_Flambeau_Flowage"
  #   )
  #   ) %>%
  plot(type = "phylogram",
     tip.color = repu_mono_MN_shades2, 
     use.edge.length = TRUE,
     node.pos = 1, 
     show.tip.label = TRUE, 
     show.node.label = T,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
nodelabels()
tiplabels()


#Test node rotations here
WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei, 42)
WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei_rotate, 41)
# WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei_rotate, 37)

#check the tip.label names
WAE_genpop_nei$tip.label
#Specify tip order (bottom to top) of the tip.labels in rotateConstr() in conjunction with node rotation to get desired format
WAE_genpop_nei_rotate %>% 
  rotateConstr(constraint = c(
    "WolfR_",
    "Lake_Wisconsin",
    "Willow_Flowage", 
    "Big_Arbor_Vitae",
    "Manitowish_Lake"



    )
  ) %>%
  plot(type = "phylogram",
     tip.color = repu_mono_MN_shades2, 
     use.edge.length = TRUE,
     node.pos = 1, 
     show.tip.label = TRUE, 
     show.node.label = T,
     edge.color = "black", 
     edge.width = 1, 
     edge.lty = 1, 
     font = 2,
     cex = 2.5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)

dev.off()
```
#Output jpg for slides
```{r}
jpeg("./IAGLR_tree.jpg", width = 32, height = 32, units = "in", quality = 100, res = 300)
WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei, 42)
WAE_genpop_nei_rotate = dendextend::rotate(WAE_genpop_nei_rotate, 41)

WAE_genpop_nei$tip.label

WAE_genpop_nei_rotate %>% 
  rotateConstr(constraint = c(
    "WolfR_",
    "Lake_Wisconsin",
    "Willow_Flowage", 
    "Big_Arbor_Vitae",
    "Manitowish_Lake"



    )
  ) %>%
  plot(type = "phylogram",
     tip.color = repu_mono_MN_shades2, 
     use.edge.length = TRUE,
     node.pos = 1, 
     show.tip.label = TRUE, 
     show.node.label = F,
     edge.color = "black", 
     edge.width = 5, 
     edge.lty = 1, 
     font = 2,
     cex = 5, 
     adj = NULL, 
     srt = 0, 
     no.margin = FALSE,
     root.edge = FALSE, 
     label.offset = 0, 
     underscore = FALSE,
     x.lim = NULL, 
     y.lim = NULL, 
     direction = "rightwards",
     lab4ut = NULL, 
     plot = TRUE,
     rotate.tree = 0, 
     open.angle = 0, 
     node.depth = 15,
     align.tip.label = FALSE)
nodelabels(WAE_genpop_nei_rotate$node.label, adj = c(1.3, -0.5), frame = "n", cex = 3,
           font = 3, xpd = TRUE)
dev.off()
```


